Automated topic naming to support analysis of software maintenance activities

### 2010-08-19 icse/icse2011.tex
10 pages or less
generating the data - edit para 1
updated ROC definition/cite; updated unsupervised methodology; table placement

\documentclass[]{sig-alternate}
 \usepackage{booktabs}
\usepackage[caption=false]{subfig}
 \usepackage[colorinlistoftodos,backgroundcolor=yellow]{todonotes}
 \usepackage{color}

@@

% Maybe specify the hard work we did annotating the data

@@

Topics in software engineering practice refer to specific shared threads of discussions between project stakeholders. These topics can include particular requirements, maintenance requests, bug fixes or other software engineering tasks. %NEIL: define topic specifically as akin to email 'thread'. It sounded too vague. Also changed from 'developers' to 'stakeholders' because developers sounds too coding oriented. Our work could apply to testers, users, etc.
While addressing a topic of development, stakeholders often produce or modify multiple development artifacts.
%
These artifacts can include source code, bug reports, revisions to source code, and test cases. 
%
Topics are typically latent, i.e., not explicitly named, and can be in several potential states: they may not be immediately resolved, they may re-occur, yet others are only dealt with briefly  (e.g., a bug report that is closed).
% semi-automatically sounds too uncertain. I think the introduction can be more muscular and bold. Don't give the reviewer any room to start doubting.

@@

% topics are numerous?
Deriving such topics is difficult. 
Topic analysis produces raw topics that are represented as unlabelled context-free word-lists or word distributions.
Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. 
This manual approach is impractical when one has to handle hundreds of different topics. To scale to meaningfully sized projects, we need automatic assistance to determine the topic label.

@@

% harp on data mining more
In general the data mining of software artifacts tends to be very project specific, yet NFRs are not. 
NFRs are prevalent in almost all software systems.
In this paper we will show that NFR related topics occur quite frequently.
This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering, specifically NFRs. In particular, we use software quality models to generalize topics across projects. 
This allows us to data mine software systems with the intention of comparing their NFR-related development topics.
% in order to achieve generalizable, cross project data mining results.

%% RESEARCH DIDNT DO THIS AND WE ARE TOTALLY AWESOME AND WE DID
% it is ok to repeat this here because it is clear and bullet pointed and the reviewers wont have any bloody excuse to not see it
In this paper, we describe \textbf{Nomen}, which is a tool for \emph{labelled topic extraction}. \textbf{Nomen} addresses two gaps in the topic mining literature:
\begin{enumerate}
  \item Topic mining of software has been limited to one project at a time. This is because traditional topic mining techniques are specific to a particular data-set. \textbf{Nomen} allows for comparisons \textit{between} projects. 
  \item Topic modeling creates word lists that require interpretation by the user to assign meaning. Like (1), this means
that it is difficult to discuss results independent of the project context. Our technique automatically assigns labels agnostic of project.
\end{enumerate}

This paper begins by introducing labelled topic extraction (Section \ref{sec:wordlist}). We show that these labels with their topics can be learned and used to classify other data-sets. We then present visualizations of named topics and their trends over time to aid communication and analysis. We use an exploratory case study of two open source database systems to show how named topics can be compared between projects.

@@

\todo[inline]{We still haven't shown that topics are useful -- a reviewer comment}

\section{Study design and execution}
Figure \ref{fig:process} gives an outline of the methodology followed. We began by gathering source data;
then we manually annotated the topics of these sources by their related NFRs;
 then applied unsupervised and supervised learning techniques to identify topics. We then use these topics to analyze the role of non-functional requirements in software maintenance.

@@

\subsection{Generating the data}
\label{sec:wordlist}
To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. We selected systems from the same application domain, so the functional requirements would be broadly similar. We selected MySQL and MaxDB as they were open-source, partially-commercial database systems. MaxDB started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project has 940 thousand lines of C source code~\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}. The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. MySQL contains 320 thousand lines of C and C++ source code.  Choosing an older version of these projects allows us to focus on projects which have moved into the maintenance phase of the software life-cycle.

@@

From that data-set, we created an XML file which separated commits into 30 day periods. This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze. For this set of periods, we created `topics' using Latent Dirichlet Allocation, a topic identification algorithm.

Typically a topic analysis tool like LDA will try to find $N$ independent word distributions within the word distributions of all the messages. Linear combinations of these $N$ word distributions are then meant to be able to recreate the word distributions of all of the underlying messages. These $N$ word distributions effectively form topics: cross-cutting collections of words relevant to one or more of our commit messages. LDA extracts topics in an unsupervised manner; the algorithm relies solely on the source data with no human intervention.

@@

\subsection{Generating word lists}
Topics are word distributions: lists of words ranked by frequency, which can be burdensome to interpret and hard to distinguish and understand. While the topic models themselves are generated automatically, what to make of them is less clear. For example, in our previous work~\cite{Hindle09ICSM}, as well as in Baldi et al.~\cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a keyword accordingly. E.g., for the word list \emph{``listener change remove add fire''}, Baldi et al. assign the keyword \emph{event-handling}. The labels are reasonable enough, but still require an expert in the field to determine them. We sought to automate the process of naming the topics.

@@

\begin{table*}
	\centering
\begin{tabular}{c|p{9cm}}
\toprule
\textbf{Label} & \textbf{Related terms} \\
\midrule
\emph{Maintainability} &
testability changeability analyzability stability maintain maintainable modularity modifiability understandability interdependent dependency encapsulation decentralized modular\\ \hline
\emph{Functionality} &
security compliance accuracy interoperability suitability functional practicality functionality compliant exploit certificate secured ``buffer overflow'' policy malicious trustworthy vulnerable vulnerability accurate secure vulnerability correctness accuracy\\ \hline
\emph{Portability} &
conformance adaptability replaceability installability portable movableness movability portability specification migration standardized l10n localization i18n internationalization documentation interoperability transferability\\ \hline
\emph{Efficiency} &
``resource behaviour'' ``time behaviour'' efficient efficiency performance profiled optimize sluggish factor penalty slower faster slow fast optimization\\ \hline
\emph{Usability} &
operability understandability learnability useable usable serviceable usefulness utility useableness usableness serviceableness serviceability usability gui accessibility menu configure convention standard feature focus ui mouse icons ugly dialog guidelines click default human convention friendly user screen interface flexibility\\ \hline
\emph{Reliability} &
``fault tolerance'' recoverability maturity reliable dependable responsibleness responsibility reliableness reliability dependableness dependability resilience integrity stability stable crash bug fails redundancy error failure\\ 
\bottomrule
\end{tabular}
	\caption{NFRs and associated signifiers -- \textsf{exp2}}
	\label{tbl:wnsig}

@@

\todo[inline]{There should be a section here which is about manual annotation. It is really important and it is required by both analyses. Plus we can tell them how.}

\subsection{Unsupervised labelling}
% word list similarity
Using our three word lists, we labeled our topics where there was a match between the list and the topic. In general, this approach did not work out well as common labels dominated the less common labels. The related words for \emph{correctness}, for example, tended to be too lengthy and non-specific. 
Table \ref{tbl:wordlist} lists broad labeling results for MaxDB as an example. A \emph{named topic} is a topic with a matching label. There are \{periods X 20\} topics per project as we told LDA to extract 20 topics per period. All experiments were run on MaxDB 7.500 and MySQL 3.23 data.

%\todo[inline,color=green]{this table should have mysql}
\begin{table}

@@

\bottomrule
\end{tabular}
	\caption{Automatic topic labelling for MaxDB and MySQL}
	\label{tbl:wordlist}

@@

For \textsf{exp1}, our best performing labels (the labels matched with the most topics) were \emph{correctness} (182 topics) and \emph{testability} (121). We did not get good results for \emph{usability} or \emph{accuracy}, which were associated with fewer than ten topics. We also looked for correlation between our labels: Excluding double matches (self-correlation), our highest co-occurring terms were \emph{verifiability} and \emph{traceability}, and \emph{testability} and \emph{correctness} (76 and 62 matches, respectively).

For \textsf{exp2}, there are many more unnamed topics. Only \emph{reliability} produces a lot of matches, mostly with the word `error'. Co-occurrence results were poor.

For \textsf{exp3}, we had many more named topics. As we mentioned, the word-lists are quite broad, so there are likely to be false-positives. See the following sections for our error analysis. We found a high of 265 topics for \emph{usability}, with a low of 44 topics for \emph{maintainability}. Common co-occurrences were \emph{reliability} and \emph{usability}, \emph{efficiency} and \emph{reliability}, and \emph{efficiency} and \emph{usability} (200, 190, and 150 topics in common, respectively).

\subsubsection{Analysis of the unsupervised labelling}

\label{sec:unsuplabelling}

Based on the labels, and our manual topic labelling, we compared the results of the unsupervised word matching approach. For each quality we tried to assess whether the manual tag matched the unsupervised label assigned. Figure \ref{fig:maxdb-unsup-results} shows our results for MaxDB and MySQL. Our error is measured using the ROC area-under-curve value. This value is the area under the Receiver Operating Characteristic (\emph{ROC}) curve, sometimes referred to as \emph{AUC}. ROC values reflect a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data. A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data). ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false positive rate (1 - specificity). A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision.

\begin{figure*}
  \centering
 \includegraphics[width=\textwidth]{figures/unsupervised-bar}
 \caption{ROC values for each NFR, per project and wordlist.}
%\todo[inline]{this is too tiny in print}
  \label{fig:maxdb-unsup-results}
\end{figure*}

@@

\subsection{Supervised labelling}
\label{sec:suplearn}
Supervised labelling requires expert analysis of the correct word to assign to a topic. %In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set. 
 For MySQL 3.23 and MaxDB 7.500, we manually annotated each extracted topic in each period with the same NFR labels as \textsf{exp2} (software qualities). We looked at each period's topics, and assessed what the data -- consisting of the frequency-weighted word lists and messages -- suggested was the topic for that period. We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to this topic. For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}. We compared against this data-set, but we also used it for our supervised learning based topic classification. A topic may have more than one matching keyword.
\todo[inline]{inter-rater reliability }

@@

\section{NFRs and evolution} 
\todo[inline]{talk about this topic in one section}
\todo[inline]{reference project name Nomen more}
Using the classification results from the previous sections, we evaluated two research questions:
\begin{enumerate}
\item Do label frequencies change over time? That is, is a certain quality of more interest at one point in the life-cycle than some other? 
\item  Do the different projects differ in their relative topic interest? That is, is a particular quality more important to one project than the other projects?  
\end{enumerate}

@@

\section{NFRs between projects}

@@

\section{Limitations}
\subsection{Effectiveness}

%{Discuss how these are CHEAP techniques and although somewhat %inaccurate they don't require a tonne of work}

@@

\section{Related work}
The idea of extracting higher-level `concerns' (also known as `concepts', `aspects' or `requirements') has been approached in two ways.

>>>>

\documentclass[]{sig-alternate}
 \usepackage{booktabs}
\usepackage[caption=false]{subfig}
 \usepackage[disable,colorinlistoftodos,backgroundcolor=yellow]{todonotes}
 \usepackage{color}

@@

@@

Topics in software engineering practice are specific shared discussions between project participants. These topics can include particular requirements, maintenance requests, bug fixes or other software engineering tasks. %NEIL: define topic specifically as akin to email 'thread'. It sounded too vague. Also changed from 'developers' to 'participants' because developers sounds too coding oriented. Our work could apply to testers, users, etc.
In addressing a topic of development participants often produce or modify multiple development artifacts.
%
These artifacts can include source code, bug reports, revisions to source code, and test cases. 
%
Topics are typically latent, i.e., not explicitly named, and can be in several potential states: they may not be immediately resolved, they may re-occur, yet others are only dealt with briefly  (e.g., a bug report that is closed).
% semi-automatically sounds too uncertain. I think the introduction can be more muscular and bold. Don't give the reviewer any room to start doubting.

@@

% topics are numerous?
Deriving such topics is difficult. 
Raw topics from topic analysis are represented as unlabelled context-free word-lists or word distributions.
Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. 
This manual approach is impractical when one has to handle hundreds of different topics. To scale to meaningfully sized projects, we need automatic assistance to determine the topic label.

@@

% harp on data mining more
In general the data mining of software artifacts tends to be very project specific, yet NFRs are not. 
NFRs are prevalent in almost all software systems.
% In this paper we will show that NFR related topics occur quite frequently.
This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering, specifically NFRs. In particular, we use software quality models to generalize topics across projects. 
This allows us to data mine software systems with the intention of comparing their NFR-related development topics.
% in order to achieve generalizable, cross project data mining results.

%% RESEARCH DIDNT DO THIS AND WE ARE TOTALLY AWESOME AND WE DID
% it is ok to repeat this here because it is clear and bullet pointed and the reviewers wont have any bloody excuse to not see it
In this paper, we describe \textbf{Nomen}, which is a tool for \emph{labelled topic extraction}. \textbf{Nomen} addresses two gaps in the topic mining literature:
\begin{enumerate}
  \item Topic mining of software has been limited to one project at a time. This is because traditional topic mining techniques are specific to a particular data-set. \textbf{Nomen} allows for comparisons \textit{between} projects. 
  \item Topic modeling creates word lists that require interpretation by the user to assign meaning. Like (1), this means
that it is difficult to discuss results independent of the project context. Our technique automatically assigns labels across projects.
\end{enumerate}

This paper begins by introducing labelled topic extraction (Section \ref{sec:wordlist}). We show that these labels with their topics can be learned and used to classify other data-sets, either without training (Section \ref{sec:unsuplabelling}), or with training (Section \ref{sec:suplearn}). We then present visualizations of named topics and their trends over time to aid communication and analysis. We use an exploratory case study of two open source database systems to show how named topics can be compared between projects  (Section \ref{sec:analysis}). The paper concludes with a discussion of limitations (Section \ref{sec:limit}) and related work (Section \ref{sec:related}).

@@

\section{Study design and execution}
Figure \ref{fig:process} gives an outline of the methodology followed. We began by gathering source data. We then applied unsupervised and supervised learning techniques to identify topics. These topics were used to analyze the role of non-functional requirements in software maintenance.

@@

\subsection{Generating the data}
\label{sec:wordlist}
To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. We selected systems from the same application domain, so the functional requirements would be broadly similar. We selected MySQL and MaxDB as they were open-source, partially-commercial database systems. MaxDB started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}. The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. MySQL contains 320 thousand lines of C and C++ source code.  Choosing an older version of these projects allows us to focus on projects which have moved into the maintenance phase of the software life-cycle.

@@

From that data-set, we created an XML file which separated commits into 30 day periods. This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze. For this set of periods, we created `topics' using Latent Dirichlet Allocation, a topic identification algorithm~\cite{Blei2003}.

Typically a topic analysis tool like LDA will try to find $N$ independent word distributions within the word distributions of all the messages. Linear combinations of these $N$ word distributions are then used to recreate the word distributions of the underlying messages. These $N$ word distributions effectively form topics: cross-cutting collections of words relevant to one or more of our commit messages. LDA extracts topics in an unsupervised manner; the algorithm relies solely on the source data with no human intervention.

@@

\subsection{Generating word lists}
Topics are word distributions: lists of words ranked by frequency, which can be burdensome to interpret and hard to distinguish and understand. While the topic models themselves are generated automatically, what to make of them is less clear. For example,
% in our previous work~\cite{Hindle09ICSM}, as well as -- CITED a little too often
in Baldi et al.~\cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a keyword accordingly. E.g., for the word list \emph{``listener change remove add fire''}, Baldi et al. assign the keyword \emph{event-handling}. The labels are reasonable enough, but still require an expert in the field to determine them. We sought to automate the process of naming the topics.

@@

\begin{table*}[]
	\centering
\begin{tabular}{c|p{11cm}}
\toprule
\textbf{Label} & \textbf{Related terms} \\
\midrule
\emph{Maintainability} &
testability changeability analyzability stability maintain maintainable modularity modifiability understandability interdependent dependency encapsulation decentralized modular\\ \hline
\emph{Functionality} &
security compliance accuracy interoperability suitability functional practicality functionality compliant exploit certificate secured ``buffer overflow'' policy malicious trustworthy vulnerable vulnerability accurate secure vulnerability correctness accuracy\\ \hline
\emph{Portability} &
conformance adaptability replaceability installability portable movableness movability portability specification migration standardized l10n localization i18n internationalization documentation interoperability transferability\\ \hline
\emph{Efficiency} &
``resource behaviour'' ``time behaviour'' efficient efficiency performance profiled optimize sluggish factor penalty slower faster slow fast optimization\\ \hline
\emph{Usability} &
operability understandability learnability useable usable serviceable usefulness utility useableness usableness serviceableness serviceability usability gui accessibility menu configure convention standard feature focus ui mouse icons ugly dialog guidelines click default human convention friendly user screen interface flexibility\\ \hline
\emph{Reliability} &
``fault tolerance'' recoverability maturity reliable dependable responsibleness responsibility reliableness reliability dependableness dependability resilience integrity stability stable crash bug fails redundancy error failure\\ 
\bottomrule
\end{tabular}
	\caption{NFRs and associated signifiers -- \textsf{exp2}}
	\label{tbl:wnsig}

@@

\noindent \textbf{Creating a validation corpus} -- For MySQL 3.23 and MaxDB 7.500, we manually annotated each extracted topic in each period with the same NFR labels as \textsf{exp2} (software qualities). We looked at each period's topics, and assessed what the data -- consisting of the frequency-weighted word lists and messages -- suggested was the topic for that period. We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to this topic. For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}. We compared against this data-set, but we also used it for our supervised learning based topic classification. A topic may have more than one matching keyword.
\todo[inline]{inter-rater reliability }
% \todo[inline]{There should be a section here which is about manual annotation. It is really important and it is required by both analyses. Plus we can tell them how.}

\subsection{Unsupervised labelling}
\label{sec:unsuplabelling}

% word list similarity
Using our three word lists, we labeled our topics where there was a match between a word in the list and the same word in the topic.
Table \ref{tbl:wordlist} lists broad labeling results for MaxDB as an example. A \emph{named topic} is a topic with a matching label. There are \{periods X 20\} topics per project as we told LDA to extract 20 topics per period. All experiments were run on MaxDB 7.500 and MySQL 3.23 data.

%\todo[inline,color=green]{this table should have mysql}
\begin{table}[h]

@@

\bottomrule
\end{tabular}
	\caption{Automatic topic labelling for MaxDB and MySQL}% \todo[inline]{this stupid table is before table 1}}
	\label{tbl:wordlist}

@@

For \textsf{exp1}, our best performing labels (the labels matched with the most topics) were \emph{correctness} (182 topics) and \emph{testability} (121). We did not get good results for \emph{usability} or \emph{accuracy}, which were associated with fewer than ten topics. We also looked for correlation between our labels: Excluding double matches (self-correlation), our highest co-occurring terms were \emph{verifiability} and \emph{traceability}, and \emph{testability} and \emph{correctness} (76 and 62 matches, respectively).
% See the following sections for our error analysis.

For \textsf{exp2}, there are more unnamed topics than \textsf{exp1}. Only \emph{reliability} produces a lot of matches, mostly with the word `error'. Co-occurrence results were poor. This suggests our word lists were overly restrictive.

For \textsf{exp3}, we generally labelled more topics. As we mentioned, the word-lists are quite broad, so there are likely to be false-positives (discussed below). We found a high of 265 topics for \emph{usability}, with a low of 44 topics for \emph{maintainability}. Common co-occurrences were \emph{reliability} and \emph{usability}, \emph{efficiency} and \emph{reliability}, and \emph{efficiency} and \emph{usability} (200, 190, and 150 topics in common, respectively).

\subsubsection{Analysis of the unsupervised labelling}
Based on the labels, and our manual topic labelling, we compared the results of the unsupervised word matching approach. For each quality we assessed whether the manual tag matched the unsupervised label assigned. Figure \ref{fig:maxdb-unsup-results} shows our results for MaxDB and MySQL. Our error is measured using the \emph{ROC} (Receiver Operating Characteristic) area-under-curve value~\cite{Fawcett2006861}. ROC values provide a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data. ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false positive rate (1 - specificity). A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision. A ROC result of 0.5 would be equivalent to a random learner (that is, issuing as many false positives as true positives). The AUC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen negative instance.

\todo[inline]{since ROC is new-sounding, I wonder if a more lengthy explanation would be merited. An obvious question is why not precision/recall?}
\begin{figure*}
  \centering
 \includegraphics[width=\textwidth]{figures/unsupervised-bar}
 \caption{ROC values for each NFR, per project and wordlist.}
%\todo[inline]{this is too tiny in print}
  \label{fig:maxdb-unsup-results}
\end{figure*}

@@

\subsection{Supervised labelling}
\label{sec:suplearn}
Supervised labelling requires expert analysis of the correct word to assign to a topic. %In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set.

@@

\section{Using \textbf{Nomen} to understand software maintenance activities} 
\label{sec:analysis}
% \todo[inline]{talk about this topic in one section}
% \todo[inline]{reference project name Nomen more}
Using the classification results from the previous sections, we evaluated two research questions:
\begin{enumerate}
\item Do label frequencies change over time? That is, is a certain quality of more interest at one point in the life-cycle than some other? 
\item  Do the different projects differ in their relative topic interest? That is, is a particular quality more important to one project than the other projects?  
\end{enumerate}

@@

@@

\section{Limitations}
\label{sec:limit}
\subsection{Effectiveness}

%{Discuss how these are CHEAP techniques and although somewhat %inaccurate they don't require a tonne of work}
The unsupervised labelling had difficulty distinguishing between common labels and infrequent labels. The related words for \emph{correctness}, for example, tended to be too lengthy and non-specific. 
\todo[inline,color=red]{Abram can you explain what you mean here?}

@@

\section{Related work}
\label{sec:related}
The idea of extracting higher-level `concerns' (also known as `concepts', `aspects' or `requirements') has been approached in two ways.
